---
title: "Exploratory data analysis for Homesite Quote Conversion"
author: "Juan L. Kehoe"
date: "1/30/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

# Basic information 
Competition link: https://www.kaggle.com/c/homesite-quote-conversion    
Data download link: https://www.kaggle.com/c/homesite-quote-conversion/data  

## Purpose of the project
Predict which customer will purchase a quoted insurance plan.

## Data provided:
This dataset represents the activity of a large number of customers who are interested in buying policies from Homesite. Each QuoteNumber corresponds to a potential customer and the QuoteConversion_Flag indicates whether the customer purchased a policy.

The provided features are anonymized and provide a rich representation of the prospective customer and policy. They include specific coverage information, sales information, personal information, property information, and geographic information. Your task is to predict QuoteConversion_Flag for each QuoteNumber in the test set.   

### Files
- train.csv - the training set, contains QuoteConversion_Flag 
- test.csv - the test set, does not contain QuoteConversion_Flag   

# Exploratory data analysis
## read data and check details
```{r}
# read data
train <- read.csv("train.csv", stringsAsFactors = FALSE)
# check data dimensions
dim(train)
```
#### Types of Variable
- Row indicator: QuoteNumber   
- Target variable: QuoteConversion_Flag   
- Predictor variable: The rest of the columns   

#### Data Type
- Character: Original_Quote_Date, Field6, Field12, CoverageField8, CoverageField9, PersonalField7, PersonalField16, PersonalField17, PersonalField18, PersonalField19, PropertyField3, PropertyField4, PropertyField5, PropertyField7, PropertyField14, PropertyField28, PropertyField30, PropertyField31, PropertyField32, PropertyField33, PropertyField34, PropertyField36, PropertyField37, PropertyField38, GeographicField63, GeographicField64   
- Numeric: the rest   

## transform data type
```{r}
# transform Original_Quote_Date to Date type
train$Original_Quote_Date <- as.Date(train$Original_Quote_Date)
# extract the month information
train$Month <- as.numeric(format(train$Original_Quote_Date, "%m"))
# extract year information
train$Year <- as.numeric(format(train$Original_Quote_Date, "%Y"))
# extract day information
train$Day <- as.numeric(format(train$Original_Quote_Date, "%d"))
```
## save preprocessed data
```{r}
# save(train, test, file = "Homesite.RData")
```
## explore data in train set  
### Visulize data in train
```{r}
# delete the QuoteNumber and Original_Quote_Date variables
train$QuoteNumber = NULL
train$Original_Quote_Date = NULL
# plot for all the numeric varialbes 
library(purrr)
library(tidyr)
library(ggplot2)
library(trelliscopejs)
library(gapminder)
train %>%
  keep(is.numeric) %>% 
  gather(key=variable, value=value, -QuoteConversion_Flag)%>% 
  ggplot(aes(x=value,fill= factor(QuoteConversion_Flag), color=factor(QuoteConversion_Flag))) +
  facet_trelliscope(~ variable, scales = "free", nrow=2, ncol=3, name = "numerical", path = "pic") +
  geom_histogram(aes(y=..density..),alpha=0.5, position="identity") 

# plot all the character variables
train$QuoteConversion_Flag <- as.character(train$QuoteConversion_Flag)
library(reshape2)
train %>%
  keep(is.character) %>%
  melt(id.vars="QuoteConversion_Flag") %>%
  ggplot(aes(x=value, fill=QuoteConversion_Flag, color=QuoteConversion_Flag)) + 
  facet_trelliscope(~ variable, scales = "free", nrow=2, ncol=3, name = "categorical", path = "pic") +
  geom_bar(alpha=0.5, position=position_dodge())
 
```